### 2024-04-15 docs/projects/datavis/ggplot/comparing.qmd
ggplot - rerendered

::: {.printedClojure}
```clojure
{:layers
 [{:aes_params [],
   :stat {:compute_layer :ggproto-method, :super :ggproto-method},
   :show.legend [nil],
   :mapping nil,
   :super :ggproto-method,
   :inherit.aes [true],
   :geom_params {:na.rm [false]},
   :geom
   {:non_missing_aes ["size" "shape" "colour"],
    :draw_key :ggproto-method,
    :default_aes
    {:shape [19.0],
     :colour ["black"],
     :size [1.5],
     :fill [nil],
     :alpha [nil],
     :stroke [0.5]},
    :super :ggproto-method,
    :required_aes ["x" "y"],
    :draw_panel :ggproto-method},
   :stat_params {:na.rm [false]},
   :constructor [geom_point],
   :position {:compute_layer :ggproto-method, :super :ggproto-method},
   :data []}
  {:aes_params [],
   :stat
   {:extra_params ["na.rm" "orientation"],
    :super :ggproto-method,
    :compute_group :ggproto-method,
    :required_aes ["x" "y"],
    :setup_params :ggproto-method,
    :dropped_aes ["weight"]},
   :show.legend [nil],
   :mapping nil,
   :super :ggproto-method,
   :inherit.aes [true],
   :geom_params {:se [true], :na.rm [false], :orientation [nil]},
   :geom
   {:setup_params :ggproto-method,
    :super :ggproto-method,
    :draw_key :ggproto-method,
    :extra_params ["na.rm" "orientation"],
    :required_aes ["x" "y"],
    :draw_group :ggproto-method,
    :optional_aes ["ymin" "ymax"],
    :rename_size [true],
    :setup_data :ggproto-method,
    :default_aes
    {:colour ["#3366FF"],
     :fill ["grey60"],
     :linewidth [1.0],
     :linetype [1.0],
     :weight [1.0],
     :alpha [0.4]}},
   :stat_params
   {:na.rm [false],
    :method ["lm"],
    :n [80.0],
    :orientation [nil],
    :level [0.95],
    :fullrange [false],
    :se [true],
    :formula nil,
    :method.args [],
    :span [0.75]},
   :constructor {: nil, :method ["lm"]},
   :position {:compute_layer :ggproto-method, :super :ggproto-method},
   :data []}],
 :scales {:scales [], :super :ggproto-method},
 :mapping {:x [~ cty], :y [~ hwy], :colour [~ [factor cyl]]},
 :theme [],
 :coordinates
 {:expand [true],
  :clip ["on"],
  :limits {:x nil, :y nil},
  :super :ggproto-method,
  :default [true]},
 :facet
 {:params
  {:dir ["h"],
   :nrow nil,
   :as.table [true],
   :free {:x [false], :y [false]},
   :drop [true],
   :strip.position ["top"],
   :ncol nil,
   :facets {:cyl [~ cyl]},
   :labeller
   "function (labels, multi_line = TRUE) \n{\n    labels <- lapply(labels, as.character)\n    if (multi_line) {\n        labels\n    }\n    else {\n        collapse_labels_lines(labels)\n    }\n}\n<bytecode: 0x58d461871fd0>\n<environment: namespace:ggplot2>\nattr(,\"class\")\n[1] \"function\" \"labeller\"\n"},
  :shrink [true],
  :super :ggproto-method},
 :labels {:x ["cty"], :y ["hwy"], :colour ["factor(cyl)"]}}

@@

::: {.printedClojure}
```clojure
[[[:facet :params]
  :+
  {:dir ["h"],
   :nrow nil,
   :as.table [true],
   :free {:x [false], :y [false]},
   :drop [true],
   :strip.position ["top"],
   :ncol nil,
   :facets {:cyl [~cyl]},
   :labeller
   "function (labels, multi_line = TRUE) \n{\n    labels <- lapply(labels, as.character)\n    if (multi_line) {\n        labels\n    }\n    else {\n        collapse_labels_lines(labels)\n    }\n}\n<bytecode:_ 0x58d461871fd0>\n<environment:_ namespace:ggplot2>\nattr(,\"class\")\n[1] \"function\" \"labeller\"\n"}]]

>>>>

::: {.printedClojure}
```clojure
{:layers
 [{:aes_params [],
   :stat {:compute_layer :ggproto-method, :super :ggproto-method},
   :show.legend [nil],
   :mapping nil,
   :super :ggproto-method,
   :inherit.aes [true],
   :geom_params {:na.rm [false]},
   :geom
   {:non_missing_aes ["size" "shape" "colour"],
    :draw_key :ggproto-method,
    :default_aes
    {:shape [19.0],
     :colour ["black"],
     :size [1.5],
     :fill [nil],
     :alpha [nil],
     :stroke [0.5]},
    :super :ggproto-method,
    :required_aes ["x" "y"],
    :draw_panel :ggproto-method},
   :stat_params {:na.rm [false]},
   :constructor [geom_point],
   :position {:compute_layer :ggproto-method, :super :ggproto-method},
   :data []}
  {:aes_params [],
   :stat
   {:extra_params ["na.rm" "orientation"],
    :super :ggproto-method,
    :compute_group :ggproto-method,
    :required_aes ["x" "y"],
    :setup_params :ggproto-method,
    :dropped_aes ["weight"]},
   :show.legend [nil],
   :mapping nil,
   :super :ggproto-method,
   :inherit.aes [true],
   :geom_params {:se [true], :na.rm [false], :orientation [nil]},
   :geom
   {:setup_params :ggproto-method,
    :super :ggproto-method,
    :draw_key :ggproto-method,
    :extra_params ["na.rm" "orientation"],
    :required_aes ["x" "y"],
    :draw_group :ggproto-method,
    :optional_aes ["ymin" "ymax"],
    :rename_size [true],
    :setup_data :ggproto-method,
    :default_aes
    {:colour ["#3366FF"],
     :fill ["grey60"],
     :linewidth [1.0],
     :linetype [1.0],
     :weight [1.0],
     :alpha [0.4]}},
   :stat_params
   {:na.rm [false],
    :method ["lm"],
    :n [80.0],
    :orientation [nil],
    :level [0.95],
    :fullrange [false],
    :se [true],
    :formula nil,
    :method.args [],
    :span [0.75]},
   :constructor {: nil, :method ["lm"]},
   :position {:compute_layer :ggproto-method, :super :ggproto-method},
   :data []}],
 :scales {:scales [], :super :ggproto-method},
 :mapping {:x [~ cty], :y [~ hwy], :colour [~ [factor cyl]]},
 :theme [],
 :coordinates
 {:expand [true],
  :clip ["on"],
  :limits {:x nil, :y nil},
  :super :ggproto-method,
  :default [true]},
 :facet
 {:params
  {:dir ["h"],
   :nrow nil,
   :as.table [true],
   :free {:x [false], :y [false]},
   :drop [true],
   :strip.position ["top"],
   :ncol nil,
   :facets {:cyl [~ cyl]},
   :labeller
   "function (labels, multi_line = TRUE) \n{\n    labels <- lapply(labels, as.character)\n    if (multi_line) {\n        labels\n    }\n    else {\n        collapse_labels_lines(labels)\n    }\n}\n<bytecode: 0x5b75b1efa660>\n<environment: namespace:ggplot2>\nattr(,\"class\")\n[1] \"function\" \"labeller\"\n"},
  :shrink [true],
  :super :ggproto-method},
 :labels {:x ["cty"], :y ["hwy"], :colour ["factor(cyl)"]}}

@@

::: {.printedClojure}
```clojure
[[[:facet :params]
  :+
  {:dir ["h"],
   :nrow nil,
   :as.table [true],
   :free {:x [false], :y [false]},
   :drop [true],
   :strip.position ["top"],
   :ncol nil,
   :facets {:cyl [~cyl]},
   :labeller
   "function (labels, multi_line = TRUE) \n{\n    labels <- lapply(labels, as.character)\n    if (multi_line) {\n        labels\n    }\n    else {\n        collapse_labels_lines(labels)\n    }\n}\n<bytecode:_ 0x5b75b1efa660>\n<environment:_ namespace:ggplot2>\nattr(,\"class\")\n[1] \"function\" \"labeller\"\n"}]]

### 2024-04-15 docs/projects/datavis/ggplot/draft.qmd
ggplot - rerendered

::: {.printedClojure}
```clojure
{:layers
 [{:aes_params {:size [5.0]},
   :stat {:compute_layer :ggproto-method, :super :ggproto-method},
   :show.legend [nil],
   :mapping {:colour [~ cyl]},
   :super :ggproto-method,
   :inherit.aes [true],
   :geom_params {:na.rm [false]},
   :geom
   {:non_missing_aes ["size" "shape" "colour"],
    :draw_key :ggproto-method,
    :default_aes
    {:shape [19.0],
     :colour ["black"],
     :size [1.5],
     :fill [nil],
     :alpha [nil],
     :stroke [0.5]},
    :super :ggproto-method,
    :required_aes ["x" "y"],
    :draw_panel :ggproto-method},
   :stat_params {:na.rm [false]},
   :constructor {: nil, :mapping {: nil, :color cyl}, :size [5.0]},
   :position {:compute_layer :ggproto-method, :super :ggproto-method},
   :data []}],
 :scales
 {:scales
  [{:labels [],
    :scale_name ["position_c"],
    :breaks [],
    :call
    {:labels labels,
     :scale_name ["position_c"],
     :breaks breaks,
     :name name,
     :palette identity,
     :super ScaleContinuousPosition,
     :guide guide,
     :minor_breaks minor_breaks,
     :n.breaks n.breaks,
     : nil,
     :aesthetics [$ ggplot_global y_aes],
     :na.value na.value,
     :trans trans,
     :oob oob,
     :position position,
     :limits limits,
     :expand expand},
    :rescaler :ggproto-method,
    :name [],
    :palette :ggproto-method,
    :super :ggproto-method,
    :guide [],
    :minor_breaks [],
    :n.breaks nil,
    :aesthetics
    ["y"
     "ymin"
     "ymax"
     "yend"
     "yintercept"
     "ymin_final"
     "ymax_final"
     "lower"
     "middle"
     "upper"
     "y0"],
    :na.value [nil],
    :trans
    {:name ["identity"],
     :transform
     "function (x) \nx\n<bytecode: 0x58d45dd43140>\n<environment: namespace:base>\n",
     :inverse
     "function (x) \nx\n<bytecode: 0x58d45dd43140>\n<environment: namespace:base>\n",
     :breaks
     "function (x, n = n_default) \n{\n    x <- x[is.finite(x)]\n    if (length(x) == 0) {\n        return(numeric())\n    }\n    rng <- range(x)\n    labeling::extended(rng[1], rng[2], n, ...)\n}\n<bytecode: 0x58d461ca4718>\n<environment: 0x58d46d4d35d8>\n",
     :minor_breaks
     "function (b, limits, n) \n{\n    b <- b[!is.na(b)]\n    if (length(b) < 2) {\n        return()\n    }\n    bd <- diff(b)[1]\n    if (!reverse) {\n        if (min(limits) < min(b)) \n            b <- c(b[1] - bd, b)\n        if (max(limits) > max(b)) \n            b <- c(b, b[length(b)] + bd)\n    }\n    else {\n        if (max(limits) > max(b)) \n            b <- c(b[1] - bd, b)\n        if (min(limits) < min(b)) \n            b <- c(b, b[length(b)] + bd)\n    }\n    seq_between <- function(a, b) {\n        seq(a, b, length.out = n + 1)[-(n + 1)]\n    }\n    breaks <- unlist(Map(seq_between, b[-length(b)], b[-1]))\n    breaks <- c(breaks, b[length(b)])\n    breaks\n}\n<bytecode: 0x58d461caa018>\n<environment: 0x58d46d4d3760>\n",
     :format
     "function (x) \n{\n    if (!is.null(names(x))) {\n        return(names(x))\n    }\n    ret <- format(x, ..., trim = TRUE, justify = \"left\")\n    ret[is.na(x)] <- NA\n    ret\n}\n<bytecode: 0x58d461cad640>\n<environment: 0x58d46d4d38b0>\n",
     :domain [##-Inf ##Inf]},
    :oob :ggproto-method,
    :position ["left"],
    :limits [20.0 30.0],
    :expand [],
    :range {:super :ggproto-method}}],
  :super :ggproto-method},
 :mapping {:x [~ cty], :y [~ hwy]},
 :theme [],
 :coordinates
 {:expand [true],
  :clip ["on"],
  :limits {:x nil, :y nil},
  :super :ggproto-method,
  :default [true]},
 :facet {:shrink [true], :super :ggproto-method},
 :labels {:x ["cty"], :y ["hwy"], :colour ["cyl"]}}

@@

     "function (x) \nlog(x, base)\n<bytecode: 0x58d467489ac8>\n<environment: 0x58d46f8464a8>\n",

@@

     "function (x) \nbase^x\n<bytecode: 0x58d467489c18>\n<environment: 0x58d46f8464a8>\n",

@@

     "function (x, n = n_default) \n{\n    raw_rng <- suppressWarnings(range(x, na.rm = TRUE))\n    if (any(!is.finite(raw_rng))) {\n        return(numeric())\n    }\n    rng <- log(raw_rng, base = base)\n    min <- floor(rng[1])\n    max <- ceiling(rng[2])\n    if (max == min) {\n        return(base^min)\n    }\n    by <- floor((max - min)/n) + 1\n    breaks <- base^seq(min, max, by = by)\n    relevant_breaks <- base^rng[1] <= breaks & breaks <= base^rng[2]\n    if (sum(relevant_breaks) >= (n - 2)) {\n        return(breaks)\n    }\n    while (by > 1) {\n        by <- by - 1\n        breaks <- base^seq(min, max, by = by)\n        relevant_breaks <- base^rng[1] <= breaks & breaks <= \n            base^rng[2]\n        if (sum(relevant_breaks) >= (n - 2)) {\n            return(breaks)\n        }\n    }\n    log_sub_breaks(rng, n = n, base = base)\n}\n<bytecode: 0x58d4673d5710>\n<environment: 0x58d46fe22dc8>\n",

@@

     "function (b, limits, n) \n{\n    b <- b[!is.na(b)]\n    if (length(b) < 2) {\n        return()\n    }\n    bd <- diff(b)[1]\n    if (!reverse) {\n        if (min(limits) < min(b)) \n            b <- c(b[1] - bd, b)\n        if (max(limits) > max(b)) \n            b <- c(b, b[length(b)] + bd)\n    }\n    else {\n        if (max(limits) > max(b)) \n            b <- c(b[1] - bd, b)\n        if (min(limits) < min(b)) \n            b <- c(b, b[length(b)] + bd)\n    }\n    seq_between <- function(a, b) {\n        seq(a, b, length.out = n + 1)[-(n + 1)]\n    }\n    breaks <- unlist(Map(seq_between, b[-length(b)], b[-1]))\n    breaks <- c(breaks, b[length(b)])\n    breaks\n}\n<bytecode: 0x58d461caa018>\n<environment: 0x58d46fe22a10>\n",

@@

     "function (x) \n{\n    if (!is.null(names(x))) {\n        return(names(x))\n    }\n    ret <- format(x, ..., trim = TRUE, justify = \"left\")\n    ret[is.na(x)] <- NA\n    ret\n}\n<bytecode: 0x58d461cad640>\n<environment: 0x58d46fe228c0>\n",

>>>>

::: {.printedClojure}
```clojure
{:layers
 [{:aes_params {:size [5.0]},
   :stat {:compute_layer :ggproto-method, :super :ggproto-method},
   :show.legend [nil],
   :mapping {:colour [~ cyl]},
   :super :ggproto-method,
   :inherit.aes [true],
   :geom_params {:na.rm [false]},
   :geom
   {:non_missing_aes ["size" "shape" "colour"],
    :draw_key :ggproto-method,
    :default_aes
    {:shape [19.0],
     :colour ["black"],
     :size [1.5],
     :fill [nil],
     :alpha [nil],
     :stroke [0.5]},
    :super :ggproto-method,
    :required_aes ["x" "y"],
    :draw_panel :ggproto-method},
   :stat_params {:na.rm [false]},
   :constructor {: nil, :mapping {: nil, :color cyl}, :size [5.0]},
   :position {:compute_layer :ggproto-method, :super :ggproto-method},
   :data []}],
 :scales
 {:scales
  [{:labels [],
    :scale_name ["position_c"],
    :breaks [],
    :call
    {:labels labels,
     :scale_name ["position_c"],
     :breaks breaks,
     :name name,
     :palette identity,
     :super ScaleContinuousPosition,
     :guide guide,
     :minor_breaks minor_breaks,
     :n.breaks n.breaks,
     : nil,
     :aesthetics [$ ggplot_global y_aes],
     :na.value na.value,
     :trans trans,
     :oob oob,
     :position position,
     :limits limits,
     :expand expand},
    :rescaler :ggproto-method,
    :name [],
    :palette :ggproto-method,
    :super :ggproto-method,
    :guide [],
    :minor_breaks [],
    :n.breaks nil,
    :aesthetics
    ["y"
     "ymin"
     "ymax"
     "yend"
     "yintercept"
     "ymin_final"
     "ymax_final"
     "lower"
     "middle"
     "upper"
     "y0"],
    :na.value [nil],
    :trans
    {:name ["identity"],
     :transform
     "function (x) \nx\n<bytecode: 0x5b75ae3c23a0>\n<environment: namespace:base>\n",
     :inverse
     "function (x) \nx\n<bytecode: 0x5b75ae3c23a0>\n<environment: namespace:base>\n",
     :breaks
     "function (x, n = n_default) \n{\n    x <- x[is.finite(x)]\n    if (length(x) == 0) {\n        return(numeric())\n    }\n    rng <- range(x)\n    labeling::extended(rng[1], rng[2], n, ...)\n}\n<bytecode: 0x5b75b232cda8>\n<environment: 0x5b75bd6e2898>\n",
     :minor_breaks
     "function (b, limits, n) \n{\n    b <- b[!is.na(b)]\n    if (length(b) < 2) {\n        return()\n    }\n    bd <- diff(b)[1]\n    if (!reverse) {\n        if (min(limits) < min(b)) \n            b <- c(b[1] - bd, b)\n        if (max(limits) > max(b)) \n            b <- c(b, b[length(b)] + bd)\n    }\n    else {\n        if (max(limits) > max(b)) \n            b <- c(b[1] - bd, b)\n        if (min(limits) < min(b)) \n            b <- c(b, b[length(b)] + bd)\n    }\n    seq_between <- function(a, b) {\n        seq(a, b, length.out = n + 1)[-(n + 1)]\n    }\n    breaks <- unlist(Map(seq_between, b[-length(b)], b[-1]))\n    breaks <- c(breaks, b[length(b)])\n    breaks\n}\n<bytecode: 0x5b75b23326a8>\n<environment: 0x5b75bd6e2710>\n",
     :format
     "function (x) \n{\n    if (!is.null(names(x))) {\n        return(names(x))\n    }\n    ret <- format(x, ..., trim = TRUE, justify = \"left\")\n    ret[is.na(x)] <- NA\n    ret\n}\n<bytecode: 0x5b75b2335cd0>\n<environment: 0x5b75bd6e2198>\n",
     :domain [##-Inf ##Inf]},
    :oob :ggproto-method,
    :position ["left"],
    :limits [20.0 30.0],
    :expand [],
    :range {:super :ggproto-method}}],
  :super :ggproto-method},
 :mapping {:x [~ cty], :y [~ hwy]},
 :theme [],
 :coordinates
 {:expand [true],
  :clip ["on"],
  :limits {:x nil, :y nil},
  :super :ggproto-method,
  :default [true]},
 :facet {:shrink [true], :super :ggproto-method},
 :labels {:x ["cty"], :y ["hwy"], :colour ["cyl"]}}

@@

     "function (x) \nlog(x, base)\n<bytecode: 0x5b75b7b00b30>\n<environment: 0x5b75b582d0a0>\n",

@@

     "function (x) \nbase^x\n<bytecode: 0x5b75b7b00f90>\n<environment: 0x5b75b582d0a0>\n",

@@

     "function (x, n = n_default) \n{\n    raw_rng <- suppressWarnings(range(x, na.rm = TRUE))\n    if (any(!is.finite(raw_rng))) {\n        return(numeric())\n    }\n    rng <- log(raw_rng, base = base)\n    min <- floor(rng[1])\n    max <- ceiling(rng[2])\n    if (max == min) {\n        return(base^min)\n    }\n    by <- floor((max - min)/n) + 1\n    breaks <- base^seq(min, max, by = by)\n    relevant_breaks <- base^rng[1] <= breaks & breaks <= base^rng[2]\n    if (sum(relevant_breaks) >= (n - 2)) {\n        return(breaks)\n    }\n    while (by > 1) {\n        by <- by - 1\n        breaks <- base^seq(min, max, by = by)\n        relevant_breaks <- base^rng[1] <= breaks & breaks <= \n            base^rng[2]\n        if (sum(relevant_breaks) >= (n - 2)) {\n            return(breaks)\n        }\n    }\n    log_sub_breaks(rng, n = n, base = base)\n}\n<bytecode: 0x5b75b7a46af8>\n<environment: 0x5b75b58310d8>\n",

@@

     "function (b, limits, n) \n{\n    b <- b[!is.na(b)]\n    if (length(b) < 2) {\n        return()\n    }\n    bd <- diff(b)[1]\n    if (!reverse) {\n        if (min(limits) < min(b)) \n            b <- c(b[1] - bd, b)\n        if (max(limits) > max(b)) \n            b <- c(b, b[length(b)] + bd)\n    }\n    else {\n        if (max(limits) > max(b)) \n            b <- c(b[1] - bd, b)\n        if (min(limits) < min(b)) \n            b <- c(b, b[length(b)] + bd)\n    }\n    seq_between <- function(a, b) {\n        seq(a, b, length.out = n + 1)[-(n + 1)]\n    }\n    breaks <- unlist(Map(seq_between, b[-length(b)], b[-1]))\n    breaks <- c(breaks, b[length(b)])\n    breaks\n}\n<bytecode: 0x5b75b23326a8>\n<environment: 0x5b75b5830ce8>\n",

@@

     "function (x) \n{\n    if (!is.null(names(x))) {\n        return(names(x))\n    }\n    ret <- format(x, ..., trim = TRUE, justify = \"left\")\n    ret[is.na(x)] <- NA\n    ret\n}\n<bytecode: 0x5b75b2335cd0>\n<environment: 0x5b75b5830b98>\n",

### 2024-04-15 docs/projects/datavis/ggplot/ggplotly_cont.qmd
ggplot - rerendered

# Further exploring ggplotly's client side - DRAFT

The following template was generated from the JSON file
in the HTML generated by
`ggplotly(ggplot(mpg, aes(x=hwy, y=displ, color=factor(cyl))) + geom_point() + geom_smooth(method="lm"))`
and then gradually generalizing it as a set of Clojure functions.
This is a work-in-progress draft. Some parts are still hard-coded.

@@

::: {.sourceClojure}
```clojure
(color/palette )
```
:::

::: {.printedClojure}
```clojure
[[223.0 46.0 40.0 255.0]
 [254.0 128.0 26.0 255.0]
 [233.0 191.0 53.0 255.0]
 [129.0 187.0 66.0 255.0]
 [50.0 199.0 169.0 255.0]
 [74.0 155.0 220.0 255.0]]

```
:::

::: {.sourceClojure}
```clojure
(def colors
  ["rgba(248,118,109,1)"
   "rgba(124,174,0,1)"
   "rgba(0,191,196,1)"
   "rgba(199,124,255,1)"])
```
:::

@@

::: {.sourceClojure}
```clojure
(defn factor [column]
  (vary-meta column
             assoc :gg/factor? true))
```
:::

::: {.sourceClojure}
```clojure
(defn factor? [column]
  (:gg/factor? column))
```
:::

::: {.sourceClojure}
```clojure
(let [data toydata.ggplot/mpg
      point-layers (-> data
                       (tc/add-column "factor(cyl)" #(factor (:cyl %)))
                       (tc/group-by :cyl {:result-type :as-map})
                       (->> (sort-by key)
                            (map-indexed
                             (fn [i [group-name group-data]]
                               (let [base {:x (:hwy group-data),
                                           :y (:displ group-data)
                                           :color (colors i)
                                           :name group-name
                                           :legendgroup group-name}
                                     predictions (map
                                                  (fun/linear-regressor (:hwy group-data)
                                                                        (:displ group-data))
                                                  (:hwy group-data))]
                                 [(-> base
                                      (assoc :type :point
                                             :showlegend true
                                             :y (:displ group-data)
                                             :text (-> group-data
                                                       (texts [:hwy :displ "factor(cyl)"])))
                                      layer)
                                  (-> base
                                      (assoc :type :line
                                             :showlegend false
                                             :y predictions
                                             :text (-> group-data
                                                       ;; (tc/add-column :displ predictions)
                                                       (texts [:hwy :displ "factor(cyl)"])))
                                      layer)])))
                            (apply concat)))
      xmin (-> data :hwy fun/reduce-min)
      xmax (-> data :hwy fun/reduce-max)
      ymin (-> data :displ fun/reduce-min)
      ymax (-> data :displ fun/reduce-max)
      xaxis (axis {:minval xmin
                   :maxval xmax
                   :anchor "x"
                   :title :hwy})
      yaxis (axis {:minval ymin
                   :maxval ymax
                   :anchor "x"
                   :title :displ})]
  (kind/htmlwidgets-ggplotly
   {:x
    {:config config
     :highlight highlight
     :base_url "https://plot.ly",
     :layout (layout {:xaxis xaxis
                      :yaxis yaxis})
     :data point-layers},
    :evals [],
    :jsHooks []}))
```
:::

>>>>

# Further exploring ggplotly's client side

The following template was generated from the JSON file
in the HTML generated by
`ggplotly(ggplot(mpg, aes(x=hwy, y=displ, color=factor(cyl))) + geom_point() + geom_smooth(method="lm"))`
and then gradually generalizing it as a set of Clojure functions.
This is a work-in-progress draft. Some parts are still hard-coded (e.g., colours).

@@

::: {.sourceClojure}
```clojure
(def colors
  (->> :category10
       color/palette
       (mapv (fn [[r g b a]]
               (format "rgba(%d,%d,%d,%f)"
                       (int r)
                       (int g)
                       (int b)
                       (/ a 255.0))))))
```
:::

@@

::: {.sourceClojure}
```clojure
(delay
  (let [data toydata.ggplot/mpg
        point-layers (-> data
                         (tc/add-column "factor(cyl)" #(:cyl %))
                         (tc/group-by :cyl {:result-type :as-map})
                         (->> (sort-by key)
                              (map-indexed
                               (fn [i [group-name group-data]]
                                 (let [base {:x (:hwy group-data),
                                             :y (:displ group-data)
                                             :color (colors i)
                                             :name group-name
                                             :legendgroup group-name}
                                       predictions (map
                                                    (fun/linear-regressor (:hwy group-data)
                                                                          (:displ group-data))
                                                    (:hwy group-data))]
                                   [(-> base
                                        (assoc :type :point
                                               :showlegend true
                                               :y (:displ group-data)
                                               :text (-> group-data
                                                         (texts [:hwy :displ "factor(cyl)"])))
                                        layer)
                                    (-> base
                                        (assoc :type :line
                                               :showlegend false
                                               :y predictions
                                               :text (-> group-data
                                                         ;; (tc/add-column :displ predictions)
                                                         (texts [:hwy :displ "factor(cyl)"])))
                                        layer)])))
                              (apply concat)))
        xmin (-> data :hwy fun/reduce-min)
        xmax (-> data :hwy fun/reduce-max)
        ymin (-> data :displ fun/reduce-min)
        ymax (-> data :displ fun/reduce-max)
        xaxis (axis {:minval xmin
                     :maxval xmax
                     :anchor "x"
                     :title :hwy})
        yaxis (axis {:minval ymin
                     :maxval ymax
                     :anchor "x"
                     :title :displ})]
    (kind/htmlwidgets-ggplotly
     {:x
      {:config config
       :highlight highlight
       :base_url "https://plot.ly",
       :layout (layout {:xaxis xaxis
                        :yaxis yaxis})
       :data point-layers},
      :evals [],
      :jsHooks []})))
```
:::

### 2024-04-15 docs/projects/datavis/ggplot/representing.qmd
ggplotly - rerendered
ggplot - rerendered

---
format:
  html: {toc: true, toc-depth: 4, theme: cosmo, output-file: representing.html}
code-block-background: true
include-in-header: {text: '<link rel = "icon" href = "data:," />'}

>>>>

---
format:
  html: {toc: true, toc-depth: 4, theme: cosmo, output-file: representing.html}
fontsize: 0.9em
code-block-background: true
include-in-header: {text: '<link rel = "icon" href = "data:," />'}

### 2024-04-15 docs/projects/datavis/ggplot/study_session_20240225.qmd
ggplot - rerendered

::: {.printedClojure}
```clojure
{:layers
 [{:aes_params [],
   :stat {:compute_layer :ggproto-method, :super :ggproto-method},
   :show.legend [nil],
   :mapping nil,
   :super :ggproto-method,
   :inherit.aes [true],
   :geom_params {:na.rm [false]},
   :geom
   {:non_missing_aes ["size" "shape" "colour"],
    :draw_key :ggproto-method,
    :default_aes
    {:shape [19.0],
     :colour ["black"],
     :size [1.5],
     :fill [nil],
     :alpha [nil],
     :stroke [0.5]},
    :super :ggproto-method,
    :required_aes ["x" "y"],
    :draw_panel :ggproto-method},
   :stat_params {:na.rm [false]},
   :constructor [[$ .MEM x06c5570df1b44c78]],
   :position {:compute_layer :ggproto-method, :super :ggproto-method},
   :data []}],
 :scales {:scales [], :super :ggproto-method},
 :mapping {:x [~ cty], :y [~ hwy], :colour [~ year]},
 :theme [],
 :coordinates
 {:expand [true],
  :clip ["on"],
  :limits {:x nil, :y nil},
  :super :ggproto-method,
  :default [true]},
 :facet {:shrink [true], :super :ggproto-method},
 :labels {:x ["cty"], :y ["hwy"], :colour ["year"]}}

@@

<image id="image487" width="1" height="300" xlink:href="data:image/png;base64,iVBORw0KGgoAAAANSUhEUgAAAAEAAAEsCAIAAACNMk6CAAAABmJLR0QA/wD/AP+gvaeTAAAAjklEQVQ4jaWTQRKAMAgDd3ik//R1HuOlji3SjNYjhZAEKNt+BCgQIRRIAQRSi2l5hdr79/q1fqBQVw+ErvyipqGH5eh7fNFJi1M+9bC8OT/lrTD9fNYwT/0FpvQzwVju0eObndsd/uUa7qX2NffqNE5w7j7dbCudpXbj2ehX0n3f0ot5pX9/1xkPj79W8p41NH/nSHk1sQAAAABJRU5ErkJggg=="/>

@@

  <use xlink:href="#image487"/>

@@

<g id="surface483">

>>>>

::: {.printedClojure}
```clojure
{:layers
 [{:aes_params [],
   :stat {:compute_layer :ggproto-method, :super :ggproto-method},
   :show.legend [nil],
   :mapping nil,
   :super :ggproto-method,
   :inherit.aes [true],
   :geom_params {:na.rm [false]},
   :geom
   {:non_missing_aes ["size" "shape" "colour"],
    :draw_key :ggproto-method,
    :default_aes
    {:shape [19.0],
     :colour ["black"],
     :size [1.5],
     :fill [nil],
     :alpha [nil],
     :stroke [0.5]},
    :super :ggproto-method,
    :required_aes ["x" "y"],
    :draw_panel :ggproto-method},
   :stat_params {:na.rm [false]},
   :constructor [[$ .MEM x7d06b286781f4e75]],
   :position {:compute_layer :ggproto-method, :super :ggproto-method},
   :data []}],
 :scales {:scales [], :super :ggproto-method},
 :mapping {:x [~ cty], :y [~ hwy], :colour [~ year]},
 :theme [],
 :coordinates
 {:expand [true],
  :clip ["on"],
  :limits {:x nil, :y nil},
  :super :ggproto-method,
  :default [true]},
 :facet {:shrink [true], :super :ggproto-method},
 :labels {:x ["cty"], :y ["hwy"], :colour ["year"]}}

@@

<image id="image244" width="1" height="300" xlink:href="data:image/png;base64,iVBORw0KGgoAAAANSUhEUgAAAAEAAAEsCAIAAACNMk6CAAAABmJLR0QA/wD/AP+gvaeTAAAAjklEQVQ4jaWTQRKAMAgDd3ik//R1HuOlji3SjNYjhZAEKNt+BCgQIRRIAQRSi2l5hdr79/q1fqBQVw+ErvyipqGH5eh7fNFJi1M+9bC8OT/lrTD9fNYwT/0FpvQzwVju0eObndsd/uUa7qX2NffqNE5w7j7dbCudpXbj2ehX0n3f0ot5pX9/1xkPj79W8p41NH/nSHk1sQAAAABJRU5ErkJggg=="/>

@@

  <use xlink:href="#image244"/>

@@

<g id="surface240">
